Appendix 3: Example of a spatially varying effect of climate on citizen science count data with sdmTMB

### 2022-01-08 doc/appendix-snow-sdmTMB.Rmd
owl appendix heading changes and minor text edits

# Model fitting with `sdmTMB`

First, we will try fitting GLMMs (generalized linear mixed effects models) with a few different distributions appropriate for count data. 
<!-- We will use `normalize = TRUE` to speed up optimization.  -->
<!-- Setting `silent = FALSE` would give us optimization details as the model fits but clutter the R Markdown output. -->
Our only fixed effect will be the NAO climate index (`nao`), but because of the wide spatial extent of these surveys, we also expect that the effect of NAO might not be consistent throughout North America. 
Therefore, effect of NAO will also be allowed to vary spatially using this formula: `spatial_varying = ~ 0 + nao`.
Because there does not appear to be much correlation between years, year will be included as a random factor: `(1 | year_f)`. 
We will also include a fixed spatial random field (`spatial = "on"`) to account for consistent differences in habitat, proximity to the breeding grounds, and effects of flyways and barriers like coasts, and independent and identically distributed spatiotemporal random field (`spatiotemporal = "iid"`) to allow for annual deviations in these spatial effects due to factors like local weather and prey abundance. 
The random intercept for year allows the mean estimate for each of these spatiotemporal fields to vary. 
If years in sequence were strongly correlated with each other, we could use a random walk to link their mean estimates by setting `time_varying = ~ 1`.
It might also be appropriate to include count ID as another random effect (`(1 | CBCID)`) to account for consistent differences in habitat and human observers between count circles. 
But since its inclusion doesn't change our main result and it slows model fitting, we will rely on the spatial random effects to link observations in the same location for this example.

@@

<!-- ## Basic inspection -->
```{r save-models, warning=FALSE, message=FALSE, echo=FALSE, cache=FALSE, eval=TRUE}
if (!rds_saved) {
  saveRDS(m0, here::here("data/snow_w_main_effect_0_150km.rds"))
  saveRDS(m1, here::here("data/snow_w_main_effect_1_150km.rds"))
  saveRDS(m2, here::here("data/snow_w_main_effect_150km.rds"))
}
```

@@

<!-- ## Checking residuals -->

>>>>

# Model structure

First, we will try fitting GLMMs (generalized linear mixed effects models) with a few different distributions appropriate for count data. 
<!-- We will use `normalize = TRUE` to speed up optimization.  -->
<!-- Setting `silent = FALSE` would give us optimization details as the model fits but clutter the R Markdown output. -->
Our only fixed effect will be the NAO climate index (`nao`), but because of the wide spatial extent of these surveys, we also expect that the effect of NAO might not be consistent throughout North America. 
Therefore, effect of NAO will also be allowed to vary spatially using this formula: `spatial_varying = ~ 0 + nao`.
Because there does not appear to be much correlation between years, year will be included as a random factor: `(1 | year_f)`. 
We will also include a fixed spatial random field (`spatial = "on"`) to account for consistent differences in habitat, proximity to the breeding grounds, and effects of flyways and barriers like coasts, and independent and identically distributed spatiotemporal random field (`spatiotemporal = "iid"`) to allow for annual deviations in these spatial effects due to factors like local weather and prey abundance. 
The random intercept for year allows the mean estimate for each of these spatiotemporal fields to vary. 
If years in sequence were strongly correlated with each other, we could use a random walk to link their mean estimates by setting `time_varying = ~ 1`.
It might also be appropriate to include count ID as another random effect `(1 | CBCID)` to account for consistent differences between count circles in habitat and human observers. 
But since its inclusion doesn't change our main result and it slows model fitting, we will rely on the spatial random effects to link observations in the same location for this example.

@@

## Basic inspection
```{r save-models, warning=FALSE, message=FALSE, echo=FALSE, cache=FALSE, eval=TRUE}
if (!rds_saved) {
  saveRDS(m0, here::here("data/snow_w_main_effect_0_150km.rds"))
  saveRDS(m1, here::here("data/snow_w_main_effect_1_150km.rds"))
  saveRDS(m2, here::here("data/snow_w_main_effect_150km.rds"))
}
```

@@

## Checking residuals
